### 2016-08-25 section/0_intro.tex
finish ab and intro

Performance bugs\footnote{We also refer performance bugs as performance problems
following previous works in this area~\cite{PerfBug,Alabama,SongOOPSLA2014}} 
are software implementation mistakes that cause unnecessary performance
degradation in software. They widely exist in deployed software due to the 
complexity of modern software and the lack of performance testing support
~\cite{s2e,PerfBug,perf.fse10,rily.perftest,perfantipattern,xiao13:context}. 
They annoy end users and waste energy during production runs, and 
have already caused highly publicized failures~\cite{ACA-health,colorado}.
Tools that can help
developers quickly and accurately diagnose performance problems
are sorely desired.

Like general failure diagnosis, 
performance diagnosis starts from studying a problem symptom and hopefully ends
at identifying the root cause and suggesting a fix strategy. 
In the context of performance problems,
the symptom is 
execution 
slowness \cite{SongOOPSLA2014}; the root cause is about
which code region is inefficient and why.
An effective diagnosis tool can help developers quickly
and correctly figure out a fix to the performance problem.

@@

\item Accuracy. 
First, the inefficient code regions need to be accurately located.
Second, the reason a specific region is inefficient needs to be accurately
explained, so that developers can fix the problem.

\item Performance. 
Diagnosis often requires collecting run-time information. The lower the overhead
is, the easier for the diagnosis tool to be deployed, especially for 
production-run usage. 
\end{itemize}

@@

Performance bug detection tools use static or dynamic analysis to identify
code regions that match specific inefficiency patterns 
\cite{Alabama,CARAMEL, Cachetor,Xu:2010:FLD:1806596.1806617,Dufour:2008:STC:1453101.1453111, Xu:2009:GFP:1542476.1542523, Xu:2010:DIC:1806596.1806616,IsilDillig.PLDI15}. 
Unfortunately, these tools are not designed for and are consequently
unsuitable for performance diagnosis.
They are not designed to provide \textit{coverage} for a wide
variety of real-world problems; their analysis is not guided by
performance failure symptom, and hence are at disadvantage in terms of diagnosis
\textit{accuracy}; dynamic detection tools often lead to 
10X slowdowns or more \cite{Cachetor,Xu:2010:FLD:1806596.1806617,Alabama}, 
not ideal in terms of \textit{performance}.

Recently, progress has been made on statistical debugging for
performance diagnosis \cite{SongOOPSLA2014}.
This approach compares runs with and without problematic performance, and
accurately identifies control-flow constructs, such as a branch $b$ or 
a loop $l$, that are most correlated with
the execution slowness.%, referred to as slowness-predictor.
Unfortunately, this approach is \textit{not} effective
for loop-related 
performance problems, which contribute to two thirds of
real-world performance problems studied in previous work 
\cite{SongOOPSLA2014,PerfBug}. It cannot
tell whether and how loop $l$ is inefficient 
and hence is not very helpful
in fixing performance problems.

@@

Figure~\ref{fig:GCC27733} shows a performance bug in GCC. 
Function \texttt{mult\_alg} is a recursive function 
that computes the best algorithm for 
multiplying \texttt{t}, a time-consuming computation. 
At run time, \texttt{mult\_alg} is often invoked for many times, and often
with the same parameter partly due to its recursive nature.
To avoid redundant computation across different instances of
\texttt{mult\_alg} with the same parameter, developers used a
hash-table
\texttt{alg\_hash} to remember which parameter \texttt{t} has been processed
in the past and what is the result. 
Unfortunately, a mistake in the type declaration of hash-table
entry \texttt{hash\_entry} makes the memoization useless for large \texttt{t}.
In many cases, a slow path is taken, when the fast path should have been taken.
This mistake does not affect software correctness, but it causes a large amount
of redundant computation\footnote{This will be considered as a
loop-redundancy problem later in this paper, as we consider
recursive functions as a special case of loops.} 
and hurts GCC performance severely,
causing hundreds of times slow down for GCC test cases.

Debugging this performance bug is challenging. According to 
the discussion forum of GCC, developers identified
\texttt{mult\_alg} as the most time-consuming function
through profilers early on. However, developers did not figure out 
whether \texttt{mult\_alg} is inefficient, which part of it is 
inefficient (\texttt{mult\_alg} is a large function with about 400 lines of 
code), and how it is inefficient, until several weeks later.

If a tool can tell developers not only \emph{which} loop or function is
responsible for execution slowness, but also \emph{why} and \emph{how}
the loop/function is inefficient, 
the diagnosis and bug fixing
process would be much easier.

Clearly, more research is needed to improve the state of the art of performance
diagnosis --- better diagnosis
coverage, better diagnosis accuracy, and low run-time
overhead for common performance problems,
especially loop-related performance problems.

\subsection{Contributions}
This paper presents a tool, \Tool, that can help effectively diagnose
inefficient loop problems, the most common type of performance problems
\cite{SongOOPSLA2014,PerfBug}, with good coverage, accuracy, and performance.

@@

First, figuring out a root-cause taxonomy for inefficient loops.
%Specifically, we consider \textit{resultless} and
%\textit{redundancy} as two main categories of root causes for inefficient
%loops, representing loops that spend effort without producing
%any results or producing redundant results. They are further divided into
%sub-categories based on the granularity of the resultless or redundant work. 
Our taxonomy categorizes all inefficient loops into two main categories:
\emph{resultless}, when a large amount of computation produces
no side effect, and \emph{redundancy}, when a large amount of 
computation produces already-available results.
Each main category is then further divided to sub-categories.
We have strived to make our taxonomy both general enough to cover
common inefficient loops, and also specific enough to guide the failure
diagnosis of \Tool, and eventually
help developers understand
and fix performance problems. More details are
in Section \ref{sec:study}.

Second, building a tool(kit) \Tool that can automatically and accurately
identify whether and how a suspicious loop is inefficient, 
together with fix-strategy suggestions. 
We achieve this following several principles:

@@

\item Taxonomy guided design. To provide good coverage, we follow
the root-cause taxonomy discussed above and design analysis routines for
every root-cause category. Given a candidate loop, \Tool
applies a series of analysis to it to see if it matches any type of inefficiency.
%We will also put the analysis results from different analysis routines together
%to figure out the most likely root cause category for an inefficient loop.

\item Static-dynamic hybrid analysis.
As we will see, static analysis alone cannot accurately identify 
inefficiency root causes, especially because some inefficiency problems only
happen under specific workload. However, pure dynamic analysis will 
cause too large 
run-time overhead. Therefore, we take a hybrid approach to achieve both
performance and accuracy goals.
\end{itemize}

Third, using sampling to further lower the run-time overhead of \Tool, without
degrading diagnosis capability. Random sampling is a natural fit for performance
diagnosis due to the repetitive nature of inefficient code, especially
inefficient loops.

>>>>

Performance bugs\footnote{We also refer performance bugs as performance problems
following previous works in this area~\cite{PerfBug,Alabama,SongOOPSLA2014}} 
are software implementation mistakes that cause unnecessary performance
degradation in software. They widely exist in deployed software due to the 
complexity of modern software
~\cite{PerfBug,perf.fse10,rily.perftest,perfantipattern,xiao13:context}. 
They annoy end users and waste energy during production runs, and 
have already caused highly publicized failures~\cite{ACA-health,colorado}.
Tools that can help
developers quickly and accurately diagnose performance problems
are sorely desired.

Performance diagnosis is different from performance testing or
bug detection: it does \textit{not} aim to expose previously unknown performance
anomalies. 
Instead, it is similar with general failure diagnosis: 
it is applied \textit{after} an unexpected software behavior (i.e., symptom)
is observed, hoping to identify the root cause of this symptom and suggest
strategies to eliminate this symptom.
In the context of performance problems, the symptom is execution 
slowness \cite{SongOOPSLA2014}; the root cause is about
which code region is inefficient and why.
An effective diagnosis tool can help developers quickly and correctly understand 
and fix already-exposed performance symptoms.

@@

\item Accuracy. 
Which code regions are inefficient and why they are inefficient
need to be accurately identified.

\item Performance. 
Diagnosis often requires collecting run-time information. The lower the overhead
is, the easier for the tool to deploy, especially for 
production-run usage. 
\end{itemize}

@@

Performance bug detection tools use program analysis to identify
code regions that match specific inefficiency patterns 
\cite{Alabama,CARAMEL, Cachetor,Xu:2010:FLD:1806596.1806617,Dufour:2008:STC:1453101.1453111, Xu:2009:GFP:1542476.1542523, Xu:2010:DIC:1806596.1806616,IsilDillig.PLDI15}. 
Unfortunately, these tools are not designed and hence are unsuitable
to identify code regions that contribute to a specific slowness symptom.
They do not provide \textit{coverage} for a wide
variety of real-world problems; they are not guided by
performance failure symptoms, and hence are at disadvantage in terms of diagnosis
\textit{accuracy}; dynamic detection tools often lead to 
10X slowdowns or more \cite{Cachetor,Xu:2010:FLD:1806596.1806617,Alabama}, 
not ideal in terms of \textit{performance}.

Recently, progress has been made on statistical debugging for
performance diagnosis \cite{SongOOPSLA2014}.
This approach compares runs with and without problematic performance, and
identifies control-flow constructs, such as a branch $b$ or 
a loop $l$, that are most correlated with
the execution slowness.
Unfortunately, this approach is \textit{not} effective
for loop-related 
performance problems, which contribute to two thirds of
real-world performance problems studied in previous work 
\cite{SongOOPSLA2014,PerfBug}. It cannot
tell whether and how loop $l$ is inefficient 
and hence is limited in its diagnosis \textit{coverage} and \textit{accuracy}.

@@

Figure~\ref{fig:GCC27733} shows a performance problem in GCC. 
Recursive function \texttt{mult\_alg} computes the best algorithm for 
multiplying \texttt{t}, a time-consuming computation. 
At run time, \texttt{mult\_alg} is often invoked for many times, and often
with the same parameter partly due to its recursive nature.
To avoid redundant computation across different instances of
\texttt{mult\_alg}, developers used a
hash-table
\texttt{alg\_hash} to remember which parameter \texttt{t} has been processed
in the past and what is the result. 
Unfortunately, a mistake in the type declaration of hash-table
entry \texttt{hash\_entry} makes the memoization useless for large \texttt{t}.
In many cases, a slow path is taken, when the fast path should have been taken.
This mistake does not affect software correctness, but causes a large amount
of redundant computation\footnote{This will be considered as a
loop-redundancy problem later in this paper, as we consider
recursive functions as a special case of loops.} 
and hurts GCC performance severely,
causing hundreds of times slow down for GCC test cases.

Debugging this performance problem is challenging. According to 
the discussion forum of GCC, developers identified
\texttt{mult\_alg} as the most time-consuming function
through profilers early on. However, they did not figure out 
whether \texttt{mult\_alg} is inefficient, which part of it is 
inefficient (\texttt{mult\_alg} is a large function with about 400 lines of 
code), and how it is inefficient, until several weeks later.

If a tool can tell developers not only \emph{which} loop or function is
responsible for execution slowness, but also \emph{why} and \emph{how}
it is inefficient, 
diagnosis and fixing
would be much easier.

\iffalse
%to save space
Clearly, more research is needed to improve the state of the art of performance
diagnosis --- better diagnosis
coverage, better diagnosis accuracy, and low run-time
overhead for common performance problems,
especially loop-related performance problems.
\fi

\subsection{Contributions}
This paper presents a tool \Tool that can help effectively diagnose
inefficient loop problems, the most common type of performance problems
\cite{SongOOPSLA2014,PerfBug}, with good coverage, accuracy, and performance.
After users or developers observe a performance failure symptom, \Tool can 
automatically judge whether the symptom is caused by inefficient loops and
provide detailed root-cause information that helps understand and fix 
inefficient loops.

@@

First, figuring out a root-cause taxonomy for inefficient loops.
%Specifically, we consider \textit{resultless} and
%\textit{redundancy} as two main categories of root causes for inefficient
%loops, representing loops that spend effort without producing
%any results or producing redundant results. They are further divided into
%sub-categories based on the granularity of the resultless or redundant work. 
Our taxonomy categorizes all inefficient loops into two main categories:
\emph{resultless}, when a large amount of computation produces
no side effects, and \emph{redundancy}, when a large amount of 
computation produces already-available results.
Each main category is further divided to sub-categories.
We strive to make the taxonomy both general enough to cover
common inefficient loops, and specific enough to guide the design
of \Tool and eventually help developers understand
and fix performance problems
(Section \ref{sec:study}).

Second, building a tool(kit) \Tool that can automatically and accurately
identify whether and how a suspicious loop is inefficient, 
following these principles (Section \ref{sec:design}):

@@

\item Taxonomy guided design. To provide good coverage, we follow
the root-cause taxonomy discussed above and design analysis routines for
every root-cause category. Given a candidate loop, \Tool
applies a series of analysis to see if it matches any type of inefficiency.
%We will also put the analysis results from different analysis routines together
%to figure out the most likely root cause category for an inefficient loop.

\item Static-dynamic hybrid analysis.
Static analysis alone cannot accurately identify 
inefficiency root causes, because some inefficiency problems only
happen or matter under specific workload. However, pure dynamic analysis will 
cause too large 
run-time overhead. Therefore, we use a hybrid approach to achieve both
performance and accuracy goals.
\end{itemize}

Third, using sampling to further lower the run-time overhead of \Tool, without
degrading diagnosis capability. Random sampling is a natural fit for performance
diagnosis due to the repetitive nature of inefficient loops (Section \ref{sec:inst}).

### 2016-08-25 section/abstract.tex
finish ab and intro

Writing efficient software is difficult.
Design and implementation defects can easily
cause severe performance 
degradation. 
Unfortunately, existing performance diagnosis techniques are still preliminary.
%failing to provide the desired diagnosis coverage, accuracy, and performance. 
Performance-bug detectors can identify specific type
of inefficient computation, but are not suitable for accurately diagnosing
a wide variety of real-world performance problems. 
Profilers can locate code regions that consume resources, but 
not the ones that \textit{waste} resources.
Statistical performance diagnosis can
identify loops or branches that are most correlated with a performance
symptom, but cannot decide whether and why a loop is inefficient, and
how developers might fix it.

In this paper, we first design a root-cause
and fix-strategy taxonomy for real-world inefficient loops, 
one of the most common performance problems in the field.
We then design a static-dynamic hybrid analysis tool, \Tool, to
provide accurate performance diagnosis for loops.
We further use sampling techniques to lower the run-time overhead without
degrading the accuracy or latency of \Tool diagnosis. 
Evaluation using real-world performance
problems shows that \Tool can provide better coverage and accuracy than
existing techniques, with low overhead.

>>>>

Writing efficient software is difficult.
Design and implementation defects can
cause severe performance 
degradation. 
Unfortunately, existing performance diagnosis techniques like profilers are still preliminary.
%failing to provide the desired diagnosis coverage, accuracy, and performance. 
%Performance-bug detectors can identify specific type
%of inefficient computation, but are not suitable for accurately diagnosing
%a wide variety of real-world performance problems. 
They can locate code regions that consume resources, but 
not the ones that \textit{waste} resources.
%Statistical performance diagnosis can
%identify loops or branches that are most correlated with a performance
%symptom, but cannot decide whether and why a loop is inefficient, and
%how developers might fix it.

In this paper, we first design a root-cause
and fix-strategy taxonomy for inefficient loops, 
one of the most common performance problems in the field.
We then design a static-dynamic hybrid analysis tool, \Tool, to
provide accurate performance diagnosis for loops.
We further use sampling techniques to lower the run-time overhead without
degrading the accuracy or latency of \Tool diagnosis. 
Evaluation using real-world performance
problems shows that \Tool can provide better coverage and accuracy than
existing techniques, with low overhead.
